Course 5 Project 1 R Markdown File
========================

This report includes the code that I used to generate the output:

1.Load data and transform dataset's date from factor to date class

```{r loadData}
data<-read.csv("activity.csv")
str(data)
data$date<-as.Date(as.character(data$date))
```

2.Histogram of the total number of steps taken each day

```{r histStepsDayNARm}
stepsDayNARm<-with(data,tapply(steps,date,FUN=sum,na.rm=TRUE))
hist(stepsDayNARm,xlab="Number of steps taken each day",main="Histogram of Number of steps taken each day")
```

3.Mean and median number of steps taken each day

```{r computeMeanMedianNARm}
meanStepsDayNARm<-mean(stepsDayNARm)
medianStepsDayNARm<-median(stepsDayNARm)
```

Mean of steps taken each day is `r meanStepsDayNARm`

Median of steps taken each day is `r medianStepsDayNARm`

4.Time series plot of the average number of steps taken

```{r meanStepsInterval}
meanStepsIntervalNARm<-with(data,tapply(steps,interval,FUN=mean,na.rm=TRUE))
meanStepsIntervalDFNARm<-data.frame(interval=names(meanStepsIntervalNARm),meanSteps=meanStepsIntervalNARm)
meanStepsIntervalDFNARm$interval<-as.numeric(as.character(meanStepsIntervalDFNARm$interval))
plot(meanStepsIntervalDFNARm$interval,meanStepsIntervalDFNARm$meanSteps,type='l',xlab="5-min interval",ylab="Avg of steps",main="Average of steps in each 5-min interval across all days")
```

5.The 5-minute interval that, on average, contains the maximum number of steps

```{r maxMeanStepsInterval}
maxMeanStepsIntervalNARm<-meanStepsIntervalDFNARm[match(max(meanStepsIntervalNARm),meanStepsIntervalNARm),1]
```

The 5-minute interval that, on average, contains the maximum number of steps is `r maxMeanStepsIntervalNARm`

6.Code to describe and show a strategy for imputing missing data

```{r missingValues}
missingValuesCount<-sum(is.na(data$steps))   # calculate number of rows with missing step value
missingIndex<-which(is.na(data$steps))   # calculate index of row with missing step value
newdata<-data   # create a new dataset that is equal to the original dataset
for(i in 1:length(missingIndex)){   # replace missing value with mean for that 5-min interval across all days
    newdata[missingIndex[i],1]<-meanStepsIntervalDFNARm[meanStepsIntervalDFNARm$interval==newdata[missingIndex[i],3],2]
}
```

The total number of missing values in the dataset is `r missingValuesCount`

The strategy for filling in all of the missing values in the dataset is to use the mean for that 5-minute interval across all days.

7.Histogram of the total number of steps taken each day after missing values are imputed

```{r histStepsDayNARpl}
stepsDayNARpl<-with(newdata,tapply(steps,date,FUN=sum))
hist(stepsDayNARpl,xlab="Number of steps taken each day",main="Histogram of Number of steps taken each day")
meanStepsDayNARpl<-mean(stepsDayNARpl)
medianStepsDayNARpl<-median(stepsDayNARpl)
```
Mean of steps taken each day after missing values are imputed is `r meanStepsDayNARpl`

Median of steps taken each day after missing values are imputed is `r medianStepsDayNARpl`

8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r wDay}
library(ggplot2)
newdata$wDay<-factor(grepl("S.+",weekdays(newdata$date)),levels=c(TRUE,FALSE),labels=c('weekend','weekday'))
ggplot(newdata,aes(y=steps,x=interval))+xlab("Interval")+ylab("Number of steps")+ggtitle("Average of steps in each 5-min interval")+stat_summary(fun.y=mean,geom="line")+facet_wrap(~wDay,nrow=2,ncol=1)
```
